\Large \bf Cyclone: Replication Middleware for NVM Applications

### 2016-09-18 paper/paper.tex
Updates.

\section{Introduction}
TBD
\section{Single Node Programming Model}
Cyclone starts with the assumption that the programmer has already built a
concurrent durable application on a non-volatile heap. A number of libraries and
filesystems make this possible today [DAX, NVML, others ?]. As a starting point,
we support applications built on top of Intel's Non Volatile Memory Library
(NVML) [nvml.io]. We describe NVML's programming model in this section to provide
adequate background for the rest of this paper. The primary feature of interest
is the ability to write code that is crash consistent. In other words code that
makes changes to the non-volatile heap is enclosed in a crash consistent
transaction such that either all of the transaction executes or none of it
executes in the face of power failures. This is critical to maintaining
consistency as the data structure, being durable, must survive across power
failures. To illustrate this idea, Figure~\ref{fig:example} shows how one might
build a persistent linked list using NVML - the key idea is to enclose the
entire process of updating the linked list within a crash consistent
transaction. This ensure that the linked list is not corrupted due to a power
failure. A second point worth noting is that NVM libraries such as NVML usually
provide a way to wrap pointers to non-volatile memory in order that such memory
may be mapped to different locations in virtual memory across restarts without
impact to the accessing application.

@@

We exploit the fact that DPDKs packet buffers allow prepending data without
physically moving the data in the packet buffer. Fig [figures/packets.bmp] shows
how the layout of a packet buffer changes at the leader during replication. The
received packets contains only the RPC call from the client. We prepend a RAFT
related header (containing the previous log index, previous log term, leader's
term and leader's commit point) to packet buffer itself. To this RAFT header we
then prepend an IPv4 header. We set IP addresses in the header to use
5-tuple steering in order to place the packet in the right queue at the
receiving NIC. This composite packet buffer is then sent to every replica. In
order to properly route the packet through switches, we allocate a small buffer
containing an ethernet header for each follower replica and \emph{chain} it in
front of the prepared packet - ensuring we do not need to change the contents
of the packet itself for every replica.

@@

In order to use the hash table with Cyclone, we simply need to to declare the
RPC call handler as the point of entry to Cyclone at the server end - shown in
Figure~\ref{fig:server_api}. The client end issues RPC calls via Cyclone's
client side library as shown in Figure~\ref{fig:client_api}. Crucially neither
side involves fault awareness or fault recovery code - a key goal for Cyclone.

@@

\begin{figure}
{ \scriptsize
\begin{verbatim}
int exec_rpc_call(void * input, int input_size, void *output)
{ 
  ...
}

struct rpc_callbacks_t callbacks = {
.exec_call = exec_rpc_call,
...
};

main()
{
  ...
  cyclone_start(callbacks, ...);
  ...
}
\end{verbatim}
}
\caption{Server side API}
\label{fig:server_api}
\end{figure}

\begin{figure}
{ \scriptsize
\begin{verbatim}
int make_rpc_call(..., void *data, int size, int rpc_number);
int get_last_rpc_number(...);
main()
{
  ...
  int rpc_number = get_last_rpc_number(...);
  while(true) {
    ...
    make_rpc_call(..., data, size, rpc_number++);

  }
  ...
}
\end{verbatim}
}
\caption{Client side API}
\label{fig:client_api}
\end{figure}

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple machines. Distributed transactions have two distinct facets: atomicity
and fault tolerance. Distributed transactions appear atomic with respect to
other atomic transactions and simple updates or reads by virtue of standard
concurrency control techniques. The most common technique is two phase locking:
cquire locks, verify that previously read objects are at the desired values,
update the locked objects and finally release the locks. We believe that NVM
application programmers - particularly those having experience with concurrent
code should not find it hard to deal with distributed concurrency. The hard part
for most programmers is likely to be the second aspect: fault tolerance. A
distributed transaction spans many machines and has a large footprint to clean
up in the event of failure. Failure handling is made more complex when
considering cascading failures - further failures while recovering from a
failure. Cyclone was designed to allow primitives such as distributed
transctions to be added without the programmer needing to worry about faults or
write fault recovery code. As an exercise we show how to add support for
distributed transactions to our hash table.

@@

We guard against fault tolerance by executing the RPC call on a quorum of
machines via Cyclone. Since execution of the RPC call leaves \emph{externally
  visible side effects} by virtue of the RPC calls it makes we require output
commit [] - the property that the execution is guaranteed to complete once it
starts. This is simply achieved by a flag in the RPC call that turns off
decoupled execution (Section~\ref{sec:decouple}). The flag causes execution to
start only after the RPC call has been successfully replicated.

@@

To ensure that the transaction is still guaranteed to complete, we use a
\emph{different} client identifier for each phase. Thus, until the transaction
completes and a new one is started the response for any RPC call continues to be
available. Figure [] shows in pseudocode the server side details of a
transaction. It is worth comparing the lack of any need for explicit fault
recovery in Cyclone with fault recovery in systems systems that explicity
require such cleanup [FARM, also refer to pages].  We believe that making
availability easy for NVM applications is key to increasing the adoption rate of
NVM by data center operators.

@@

\section{Related Work}

\section{Future Work}

>>>>

\section{Introduction}
TBD
\section{Single Node Programming Model}
Cyclone starts with the assumption that the programmer has already built a
concurrent durable application on a non-volatile heap. A number of libraries and
filesystems make this possible today [DAX, NVML, others ?]. As a starting point,
we support applications built on top of Intel's Non Volatile Memory Library
(NVML) [nvml.io]. We describe NVML's programming model in this section to
provide adequate background for the rest of this paper. The primary feature of
interest is the ability to write code that is crash consistent. In other words
code that makes changes to the non-volatile heap is enclosed in a crash
consistent transaction such that either all of the transaction executes or none
of it executes in the face of power failures. This is critical to maintaining
consistency as the data structure, being durable, must survive across power
failures. To illustrate this idea, Figure~\ref{fig:example} shows how one might
build a persistent linked list using NVML - the key idea is to enclose the
entire process of updating the linked list within a crash consistent
transaction. This ensure that the linked list is not corrupted due to a power
failure. A second point worth noting is that NVM libraries such as NVML usually
provide a way to wrap pointers to non-volatile memory (using {\tt TOID} macros)
in order that such memory may be mapped to different locations in virtual memory
across restarts without impact to the accessing application.

@@

We exploit the fact that DPDKs packet buffers allow prepending data to a packet
without physically moving the data in the packet buffer. Fig
[figures/packets.bmp] shows how the layout of a packet buffer changes at the
leader during replication. The received packets contains only the RPC call from
the client. We prepend a RAFT related header (containing the previous log index,
previous log term, leader's term and leader's commit point) to packet buffer
itself. To this RAFT header we then prepend an IPv4 header. We set IP addresses
in the header to use 5-tuple steering in order to place the packet in the right
queue at the receiving NIC. This composite packet buffer is then sent to every
replica. In order to properly route the packet through switches, we allocate a
small buffer containing an ethernet header for each follower replica and
\emph{chain} it in front of the prepared packet - ensuring we do not need to
change the contents of the packet itself for every replica.

@@

In order to use the hash table with Cyclone, we simply need to to declare the
RPC call handler as the point of entry to Cyclone at the server end.  The client
end issues RPC calls via Cyclone's client side library. Crucially neither side
involves fault awareness or fault recovery code - a key goal for Cyclone.

@@

@@

\subsection{Distributed Transactions}
A useful primitive for many distributed systems is the ability to run
distributed transactions - an atomic sequence of conditional updates spanning
multiple machines. Distributed transactions have two distinct facets: atomicity
and fault tolerance. Distributed transactions appear atomic with respect to
other atomic transactions and simple updates or reads by virtue of standard
concurrency control techniques. A common technique is strict two phase
locking: acquire all locks, update the locked objects and finally release all
the locks. We believe that NVM application programmers - particularly those
having experience with concurrent code should not find it hard to deal with
distributed concurrency. The hard part for most programmers is likely to be the
second aspect: fault tolerance. A distributed transaction spans many machines
and has a large footprint to clean up in the event of failure. Failure handling
is made more complex when considering cascading failures - further failures
while recovering from a failure. Cyclone was designed to allow primitives such
as distributed transctions to be added without the programmer needing to worry
about faults or write fault recovery code. As an exercise we show how to add
support for distributed transactions to our hash table.

@@

We guard against faults by executing the RPC call on a quorum of
machines via Cyclone. Since execution of the RPC call leaves \emph{externally
  visible side effects} by virtue of the RPC calls it makes we require output
commit [] - the property that the execution is guaranteed to complete once it
starts. This is simply achieved by a flag in the RPC call that turns off
decoupled execution (Section~\ref{sec:decouple}). The flag causes execution to
start only after the RPC call has been successfully replicated.

@@

To ensure that the transaction is still guaranteed to complete, we use a
\emph{different} client identifier for each phase. Thus, until the transaction
completes and a new one is started the response for any RPC call continues to be
available. Figure~\ref{fig:dist_tx} shows in pseudocode the server side details
of a transaction. One initiated, the distributed transaction is guaranteed to
complete given quorum availabilty. However, if the client requires intimation of
completion, the final transaction state can be written to a distinguished
partition as part of the transaction itself.

We note that this design is not intended for performance, under the assumption
that distributed transactions will be infrequent compared to single key
operations.  On the other hand, it is worth comparing the lack of any need for
explicit fault recovery in Cyclone with fault recovery in systems systems that
explicity require such cleanup [FARM, also refer to pages].  We believe that
such designs, which make availability easy for NVM applications are key to
increasing the adoption rate of NVM by data center operators.

\begin{figure}
{ \scriptsize
\begin{verbatim}
wait until RPC call is replicated
respond to client saying tx accepted for execution
acquire:
 set client identifier to 1
 for each machine
   Acquire write or read locks on the machine
   If machine reponds RPC sequence too old
     Transaction completed by someone else

update:
 set client identifier to 2
 for each machine
   Update object
   If machine reponds RPC sequence too old
     Transaction completed by someone else

release:
 set client identifier to 3
 for each machine
   Update object
   If machine reponds RPC sequence too old
     Transaction completed by someone else

optional:
 write transaction result to <quorum_tx_result>
\end{verbatim}
}
\caption{Distributed Transactions}
\label{fig:dist_tx}
\end{figure}

@@

\section{Conclusion}
